---
title: "Tut2"
author: "Mike"
date: "2024-02-27"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

A 2.1: 
a.
$$
\begin{aligned}
cdf(X_{(n)})
=&P(X_{max}<x)
\\
=&\prod_{i=1}^{n}P(X_i<x)
\end{aligned}
$$
so
$$
cdf(x) =
\begin{cases}
0 \ \  , & \text{for } x<0
\\
  (\frac{x}{\theta})^n, & \text{for } x\in [0,\theta]  \\
  1 \ \ , & \text{for } x > \theta
\end{cases}
$$ 
therefore,
$$pdf(x) =
\begin{cases}
  n(\frac{x}{\theta})^{n-1}, & \text{for } x\in (0,\theta)  \\
 0  \ \ , & \text{ o.w.}
\end{cases}
$$ 

b. obviously,$E(2\bar X)=\theta$ and
$Var(2\bar X)=\frac{4\theta^2}{3n}$ is almost zero if n is large which means
the estimator$2\bar X$ will converges in probability to $\theta$;

$E(\frac{n+1}{n}X_{(n)})=\int_0^\theta \theta(n+1)(\frac{x}{\theta})^n=\theta$ and $Var(\frac{n+1}{n}X_{(n)})=\int_0^\theta \theta^2 \frac{(n+1)^2}{n}(\frac{x}{\theta})^{n+1}-\theta^2=\frac{\theta^2}{n^2+2n}$ is almost zero if n is large which means
the estimator$2\bar X$ will converges in probability to $\theta$;
The variances of $\frac{n+1}{n}X_{(n)}$ is smaller

A 2.2:
Accoridng to  $\theta=E(\overset{\sim}{\theta})=(k_1+k_2)\theta$,to make $\overset{\sim}{\theta}$ unbaised ,we must have $k_1+k_2=1$.
$Var(\overset{\sim}{\theta})=k_1^2\sigma_1^2+k_2^2\sigma_2^2=k_1^2(\sigma_1^2+\sigma_2^2)-2k_1\sigma_2^2+\sigma_2^2$
so when $k_1=\frac{\sigma_2^2}{\sigma_1^2+\sigma_2^2}$ and $k_2=\frac{\sigma_1^2}{\sigma_1^2+\sigma_2^2}$
$\overset{\sim}{\theta}$ is unbaised and has smallest variance.


```{r}
gdp=read.csv("GDP2020.csv");
gdp[,5]=as.numeric(gdp[,5]);
gdp=na.omit(gdp);
gdp[,5]=log(gdp[,5])
gdplog=gdp[,5]
summary(gdplog)
below=quantile(gdplog,0.25)-1.5*IQR(gdplog)
above=quantile(gdplog,0.75)+1.5*IQR(gdplog)
outliers=gdp[(gdplog<below|gdplog>above),]
outliers
boxplot(gdplog)
mean(gdplog)
sd(gdplog)
```


