---
title: "tut3"
author: "Mike"
date: "2024-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1
A 2.6:
an unbaised estimator $T(X)$ has expected value $g\prime(\theta)$
CR lower bound is given by $\frac{g\prime(\theta)^2}{\mathcal{I}(\theta)}$
where $\mathcal{I}(\theta)$ is fisher information of $\theta$ is given by $$\mathcal{I}(\theta)=E[(\frac{\partial}{\partial \theta}\log f(X;\theta))^2]=-E[\frac{\partial^{2}}{\partial\theta^{2}}\log f(X;\theta)]$$
a. $g(\theta)=\theta$,for each individual independent Bernouilli trial we have 
$$\mathcal{I}(\theta)=-E[\frac{\partial^{2}}{\partial\theta^{2}}\log f(X;\theta)]=-E[\frac{\partial^{2}}{\partial\theta^{2}}(X\log\theta+(1-X)\log (1-X))]=\frac{1}{\theta(1-\theta)}$$
So for n trials, we have $\mathcal{I}(\theta)=\frac{n}{\theta(1-\theta)}$.
Therefore, CR lowerbound is $\frac{\theta(1-\theta)}{n}$
b.According to a,the only difference is $g(\theta)=\theta^2$ We can easily get CR lower bound is $\frac{4\theta^3(1-\theta)}{n}$
A 2.8
a. According to A 2.6 and the property of normal distribution,$\frac{(n-1)S^2}{\sigma^2}\sim\chi^2(n-1)$ which have expected value of $n-1$.Therefore,
$$
g(\sigma^2)=E(S^2)=\sigma^2
$$
while in one sample
$$
\mathcal{I}(\sigma^2)=-E[\frac{\partial}{\partial(\sigma^2)}(-\frac{1}{2\sigma^2}+\frac{(X-\mu)^2}{2\sigma^4})]=E[-\frac{1}{2\sigma^4}+\frac{(X-\mu)^2}{\sigma^6})]=\frac{1}{2\sigma^4}
$$
For n samples ,we have $\mathcal{I}(\sigma^2)=\frac{n}{2\sigma^4}$
So CR lower bound is $\frac{2\sigma^4}{n}$.
$$
Var(S^2)=\frac{\sigma^4Var(\chi^2(n-1))}{(n-1)^2}=\frac{2\sigma^4}{n-1}
$$
The conclusion is obvious.
b.
$$
MSE(\hat\theta)=E[(\hat\theta-\theta)^2]=Var(\hat\theta)+Bais(\hat\theta)
$$
Let $\hat{\sigma^2}=c\sum_{i=1}^n(X_i-\bar X)$
$$
MSE(\hat{\sigma^2})=Var(\hat{\sigma^2})+Bais(\hat{\sigma^2})=c^2(n-1)^2Var(S)+[c(n-1)-1]^2\sigma^4=\sigma^4[(n-1)(n+1)c^2-2(n-1)c+1]
$$
Therefore, we get min MSE when $c=\frac{1}{n+1}$.
A 2.10
$g(\theta)=\theta$,$\mathcal{I}(\theta)=\frac{n}{\sigma^2}$
Therefore, CR bound is $\frac{\sigma^2}{n}$

## 2.
$$
\int_0^1cos(2\pi x)dx=0
$$
```{r}
#numeric method
n=100
x=seq(0+1/n,1,1/n)
y=cos(x*2*pi)
intcos2pix=sum(y)/n
intcos2pix


n=1000
x=seq(0+1/n,1,1/n)
y=cos(x*2*pi)
intcos2pix=sum(y)/n
intcos2pix

```


```{r}
n=100
x=runif(n)
y=1-2*runif(n)
cos2pix=cos(2*pi*x)
sum(cos2pix)/n

n=1000
x=runif(n)
y=1-2*runif(n)
cos2pix=cos(2*pi*x)
sum(cos2pix)/n
```
```{r}
#numeric method
n=100
x=seq(0+1/n,1,1/n)
y=cos(2*pi*(x)^2)
intcos2pix2=sum(y)/n
intcos2pix2

n=1000
x=seq(0+1/n,1,1/n)
y=cos(2*pi*(x)^2)
intcos2pix2=sum(y)/n
intcos2pix2

```

```{r}
n=100
x=runif(n)
y=1-2*runif(n)
cos2pix2=cos(2*pi*x^2)
sum(cos2pix2)/n

n=1000
x=runif(n)
y=1-2*runif(n)
cos2pix2=cos(2*pi*x^2)
sum(cos2pix2)/n
```



